Theories of deliberated judgment

### 2022-08-03 simple.tex
Notation for single \phi; some minor fixes

The function ${\allleadsto}: I × \allhist → \powersetz{\Phi}$ associates to each individual and path a non-empty subset of $\Phi$, ${\allleadsto}(i, \alpha)$ representing the choice of $i$ following exposure to the arguemnts of $\alpha$, in order. The case ${\allleadsto}(i, \alpha) = \Phi$ is used to represent indifferent choice, in case this is considered observable.
If $\phi \in {\allleadsto}(i, \alpha)$, we say that $\phi$ is considered as one of the best options by $i$ considering $\alpha$. 
It is also denoted by $\alpha \ileadsto \phi$, considering ${\ileadsto} \subseteq \allhist × \Phi$ as a binary relation.
As usual with binary relations, $\hist \nileadsto \prop$ means $(\hist, \prop) \notin {\ileadsto}$ or equivalently $\phi \notin {\allleadsto}(i, \alpha)$.

@@

Such theories have the merit of simplicity, but there is a need to design more adaptive theories: in many interesting cases, no single argument will be considered decisive by all individuals in $I$. What we need is a way to argue with individuals by reacting to what they consider good arguments. The following example illustrates this need.

@@

In many decision problems, individuals will not be attracted by the same set of arguments and counter-arguments, even though they may finally opt for the same alternative. Imagine that some individual $i_1$ considers an argument $a_1$ for $\nprop$ relevant and some individual $i_2$ considers an argument $a_2$ for $\nprop$ relevant, but $i_1$ does not consider $a_2$ relevant and $i_2$ does not consider $a_1$ relevant (for example, because they already have some counter-arguments in mind to those arguments). Imagine further that there is some counter-argument $a'_1$ that can be given to $i_1$ to rebut $a_1$, and a different counter-argument $a'_2$ that can be given to $i_2$ to rebut $a_2$, so that they both have $\prop$ in their deliberated judgment when considering all these arguments.
In such situation, static theories as defined in \cref{sec:static} are ill suited: these theories need to present the same argument to all individuals and would thus need to define a bigger argument $a$ containing both $a'_1$ and $a'_2$. Such an argument could be too long to be convincing to $i_1$ or $i_2$, as it needlessly spends efforts countering an argument that the individual already considers irrelevant. This  will become an important problem if there are not only two but many a priori possible arguments $a_k$ for $\nprop$, each with their own counter-argument $a'_k$ for $\prop$.

%For a partial illustration about an ambitious setting, consider a decision to be taken about the relevance of some concrete action to help cope with climate change, and let $I$ be a set of citizens interested in the climate issue. The complete set of all arguments that will be considered relevant by at least one individual may well be very large. It may also be that, for each individual who considers relevant some arguments $A$ against the proposed concrete action, it is possible to give counter-arguments $A'$ that will convince the individual to opt for the proposal.
One of the settings discussed by citet{sunstein} permits to illustrate this point more concretely.
The authors claim that workers tend to not save enough because of psychological biases. 
It is therefore tempting to conjecture that the deliberated choice of those workers is to save more, in the sense defined here, than what their natural choice (in absence of arguments) leads to.
Let $I$ be a set of workers interested in establishing a savings plan. 
A difficulty related to this decision problem is that the complete set of all arguments that will be considered relevant by at least one individual for such a complex issue may well be very large. It may also be that, for each individual who considers relevant some arguments $A$ for saving not much (i.e., spending more in the short term), it is possible to give counter-arguments $A'$ that will convince the individual to opt for saving more.
But no single argument could be decisive for each $i \in I$: it could be that only tailored counter-arguments could convince individuals, given the complexity of the issue.
This illustrates that the deliberated choice of each individuals in $I$ may be to save more even though no static theory (\cref{def:static}) may achieve showing it, as it could be that no such theory is truthful.

The above setting illustrates the ambition of this research program, but at the price of rendering impossible in the confines of this article to define fully the set of arguments and giving a complete account of a concrete theory of deliberated choice in such a complex setting, an endeavor of its own.

>>>>

The function ${\allleadsto}: I × \allhist → \set{\phi, ¬\phi, 0}$ associates to each individual and path an element of $\Phi$ or the special symbol $0$ representing the absence of a strict preference. That is, ${\allleadsto}(i, \alpha)$ represents the choice of $i$ following exposure to the arguemnts of $\alpha$, in order. The case ${\allleadsto}(i, \alpha) = 0$ is used to represent indifference.
The notation $\alpha \ileadsto \phi$ will be used to mean ${\allleadsto}(i, \alpha) = \phi$, considering ${\ileadsto} \subseteq \allhist × \Phi$ as a binary relation.
As usual with binary relations, $\hist \nileadsto \prop$ means ${\allleadsto}(i, \alpha) ≠ \prop$ or equivalently ${\allleadsto}(i, \alpha) \in \set{\nprop, 0}$.

@@

Such theories have the merit of simplicity, but there is a need to design more adaptive theories: in many interesting cases, no single argument will be considered decisive by all individuals in $I$. What we need is a way to argue with individuals by reacting to what they consider good arguments. The following section illustrates this need.

@@

In many decision problems, individuals will not be attracted by the same set of arguments and counter-arguments, even though they may finally opt for the same alternative. Imagine that some individual $i_1$ considers an argument $a_1$ for $\nprop$ relevant and some individual $i_2$ considers an argument $a_2$ for $\nprop$ relevant, but $i_1$ does not consider $a_2$ relevant and $i_2$ does not consider $a_1$ relevant (for example, because they already have some counter-arguments in mind to those arguments). Imagine further that there is some counter-argument $a'_1$ that can be given to $i_1$ to rebut $a_1$, and a different counter-argument $a'_2$ that can be given to $i_2$ to rebut $a_2$, so that they both have $\prop$ in their deliberated judgment when considering all these arguments.
In such situation, static theories as defined in \cref{sec:static} are ill suited: these theories need to present the same argument to all individuals and would thus need to define a bigger argument $a$ containing both $a'_1$ and $a'_2$. Such an argument could be too long to be convincing to $i_1$ or $i_2$, as it needlessly spends efforts countering an argument that the individual considers irrelevant to start with. This  will become an important problem if there are not only two but many a priori possible arguments $a_k$ for $\nprop$, a small subset of which is considered relevant by a given individual, each with their own counter-argument $a'_k$ for $\prop$.

%For a partial illustration about an ambitious setting, consider a decision to be taken about the relevance of some concrete action to help cope with climate change, and let $I$ be a set of citizens interested in the climate issue. The complete set of all arguments that will be considered relevant by at least one individual may well be very large. It may also be that, for each individual who considers relevant some arguments $A$ against the proposed concrete action, it is possible to give counter-arguments $A'$ that will convince the individual to opt for the proposal.
One of the settings discussed by citet{sunstein} permits to illustrate this point more concretely.
Let $I$ be a set of workers interested in establishing a savings plan. 
The authors claim that workers tend to not save enough because of psychological biases. 
It is therefore tempting to conjecture that the deliberated choice of those workers is to save more, in the sense defined here, than what their natural choice (the one revealed in absence of arguments) leads to.
A difficulty related to this decision problem is that the complete set of all arguments that will be considered relevant by at least one individual for such a complex issue may well be very large. It may also be that, for each individual who considers relevant some arguments $A$ for saving not much (i.e., spending more in the short term), it is possible to give counter-arguments $A'$ that will convince the individual to opt for saving more.
But no single argument could be decisive for each $i \in I$: it could be that only tailored counter-arguments could convince individuals, given the complexity of the issue.
This illustrates that the deliberated choice of each individuals in $I$ may be to save more even though no static theory (\cref{def:static}) may achieve showing it, as it could be that no such theory is truthful.

%The above setting illustrates the ambition of this research program, but at the price of rendering impossible in the confines of this article to define fully the set of arguments and giving a complete account of a concrete theory of deliberated choice in such a complex setting, an endeavor of its own.
